# Base Graphics

We will continue to investigate the yeast data from earlier. Make sure that you have the data and **ggplot2** loaded into the session as part of your new script (if you started one).

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
library("ggplot2")
data(yeast, package = "jrIntroBio")
```

## Scatter plots

Let's start with a basic scatter plot of McGeoch's method `mcg` and von Heijne's method `gvh` for signal sequence recognition.

```{r label = "scatter_simple", fig.margin = TRUE, fig.cap="A basic scatter plot.", echo = TRUE}
ggplot(yeast, aes(x = mcg, y = gvh)) + 
  geom_point()
```

Whilst this can be informative at the data exploration stage, it isn't very aesthetically pleasing. First off the default axis labels are not very good.

(1) Use the `x`, `y` and `title` arguments^[Arguments are the things we pass to our function inside the `()` to control the behavior of that function.] of the `labs()` function to change the axis labels to something more sensible
    ```{r, eval = FALSE}
    ggplot(yeast, aes(x = mcg, y = gvh)) + 
      geom_point() + 
      labs(x = "McGeoch's method", 
           y = "von Heijne's method", 
           title = "Correlation of signal sequence detection measures")
    ```

(1) The range of possible values for each measure is between 0 and 1. Use the `ylim()` and `xlim()` functions to specify a new axis range. For example, for the x-axis it would be `+ xlim(0, 1)`
    ```{r, eval = FALSE}
    ggplot(yeast, aes(x = mcg, y = gvh)) + 
      geom_point() + 
      labs(x = "McGeoch's method", 
           y = "von Heijne's method") + 
      xlim(0,1) + 
      ylim(0,1)
    ```

(1) Try changing the colours of your points using the `colour` argument. You can find out what colours are allowed by name by using the `colors()` function. For instance, `colour = "red"` would change the points to red.
    ```{r, eval = FALSE}
    ggplot(yeast, aes(x = mcg, y = gvh)) + 
      geom_point(colour = "red") + 
      labs(x = "McGeoch's method", 
           y = "von Heijne's method") + 
      xlim(0,1) + 
      ylim(0,1)
    ```

(1) We could make this even neater by colouring points by a column in our data. Try replacing `colour = "red"` with `aes(colour = class)`. Why do you think we need the `aes()` function?
    ```{r, eval = FALSE}
    ggplot(yeast, aes(x = mcg, y = gvh)) + 
      geom_point(aes(colour = class)) + 
      labs(x = "McGeoch's method", 
           y = "von Heijne's method") + 
      xlim(0,1) + 
      ylim(0,1)
    
    # need the aes() function because we are mapping
    # a variable in the data set to an aesthetic
    ```

We should now have a plot that look like Figure 2. 

```{r, label = "double", fig.cap = "A scatter plot and bar chart as a target for these exercises.", echo = FALSE, fig.margin = TRUE}
ggplot(yeast, aes(x = mcg, y = gvh)) + 
  geom_point(aes(colour = class)) + 
  labs(x = "McGeoch's method", 
       y = "von Heijne's method") + 
  xlim(0,1) + 
  ylim(0,1)
```


## Extras

Here is the code for a histogram of McGeoch's method `mcg` for signal sequence recognition. 
```{r, echo = TRUE, eval = FALSE}
ggplot(yeast, aes(x = mcg)) + 
  geom_histogram()
```

(1) Try changing the binwidth using the argument `binwidth` within `geom_histogram()`. For instance, `binwidth = 0.02`
    ```{r, eval = FALSE}
    ggplot(yeast, aes(x = mcg)) + 
      geom_histogram(binwidth = 0.02)
    ```

(1) Try adding `fill = class` to the aesthetic mapping function, `aes()`. What happens?
    ```{r, eval = FALSE}
    ggplot(yeast, aes(x = mcg, fill = class)) + 
      geom_histogram(binwidth = 0.02)
    # different colour for each fill
    ```
    
(1) Try adding the argument `alpha = 0.2` to the `geom_histogram()` function. What happens?
    ```{r, eval = FALSE}
    ggplot(yeast, aes(x = mcg)) + 
      geom_histogram(binwidth = 0.02, alpha = 0.02)
    # alpha controls the transparency
    ```
    
(1) Try changing `geom_histogram()` to `geom_density()`
    ```{r, warning = FALSE, fig.width=6, fig.height=3, fig.cap = "Density plots of McGeoch's method coloured by class"}
    ggplot(yeast, aes(x = mcg, fill = class)) + 
      geom_density(binwidth = 0.02, alpha = 0.2)
    ```


<!-- (1) We saw in the notes how to create a bar chart from some categorical data by using a combination of the `table()` function and the `barplot()` function. We are going to do similar here to create a bar chart that looks at the distribution of the mpaa rating of films for both comedies and non comedies. First use `table()` on both the Comedy and mpaa variables of the movies data frame and assign your result to a variable.  -->

<!--     ```{r} -->
<!--     tab = table(movies$Comedy, movies$mpaa) -->
<!--     ``` -->


<!-- (1) We can then use `barplot()` on the result.^[Note that which ever variable is spread as columns in your `table()` results, this is the groupings column of your barplot.] -->

<!--     ```{r, eval = FALSE} -->
<!--     barplot(tab) -->
<!--     ``` -->

<!-- (1) I think this is less informative than if the bars in each group were side by side. We can make this be the case by specifying the argument `beside = TRUE` -->

<!--     ```{r, eval = FALSE} -->
<!--     barplot(tab, beside = TRUE) -->
<!--     ``` -->

<!-- (1) Hopefully you now have four groups of 2 bars along the x axis. If you haven't then try using `t()` for transpose around your table results before your barplot. Now we want the colours to match our original scatter plot, i.e black for non Comedy films and red for Comedies. This should be relatively easy here since we have already got Comedies being shown next to non Comedies in the graph. So we can just give a vector of colours to colour those bars in. Try `col = 1:2`. Do you understand why this works? -->

<!--     ```{r, eval = FALSE} -->
<!--     barplot(tab, beside = TRUE, col = 1:2) -->
<!--     ``` -->

<!-- (1) Finally make sure that your graph has meaningful axis labels and title. -->

<!--     ```{r, eval = FALSE} -->
<!--     barplot(tab, beside = TRUE, col = 1:2, -->
<!--             xlab = "MPAA Ratings", ylab = "Frequency", -->
<!--             main = "Comedy films in red") -->

<!--     ``` -->

<!-- ## Question 2 -->

<!-- Using the same yeast data: -->

<!-- (1) Make sure you have loaded **dplyr** comfortable with what the data looks like using `head()` and `colnames()`. -->

<!-- ```{r} -->
<!-- head(yeast) -->
<!-- colnames(yeast) -->
<!-- ``` -->


<!-- (1) Create a subset of extracellular proteins, i.e `class == EXC` called `exc`. -->

<!-- ```{r} -->
<!-- exc = yeast[yeast$class == "EXC",] -->
<!-- ``` -->

<!-- (1) What is the mean value for McGeoch's signal sequence method for extracellular proteins? -->

<!-- ```{r} -->
<!-- mean(exc$mcg) -->
<!-- ``` -->

<!-- (1) Are there more nuclear or mitochondrial proteins in the dataset? Hint: you can use the function `nrow()` to count the number of rows of a data.frame. -->

<!-- ```{r} -->
<!-- nuclear = yeast[yeast$class == "NUC",] -->
<!-- nrow(nuclear) -->
<!-- mito = yeast[yeast$class == "MIT",] -->
<!-- nrow(mito) -->
<!-- ## more nuclear -->
<!-- ``` -->

<!-- (1) How many proteins had a von Heijne's score of less than 0.2? -->

<!-- ```{r} -->
<!-- nrow(yeast[yeast$gvh < 0.2, ]) -->
<!-- ``` -->

